\vspace{1.0cm} \Large \bf Integrating influenza antigenic dynamics with molecular evolution \\ or \\ Cartographic characterization of the antigenic, evolutionary and competitive dynamics of seasonal influenza viruses

### 2013-02-19 flux.tex
Fix typos.

We follow-up this analysis with a more detailed analysis of year-to-year variation in antigenic drift and clade-specific incidence in the USA.
We measure year-to-year antigenic drift within each clade as described in the previous section and shown in Figure \ref{jumps}.
We measure seasonal incidence of each clade by taking average influenza-like illness (ILI) percentage and multiplying this by the proportion of viruses attributable to a clade for each season.
This measure of incidence has previously been shown to have have predictive power in the analysis of seasonal influenza trends \cite{Goldstein11}.
In correlating antigenic drift to seasonal incidence we standardize this measure of incidence across seasons and across clades to measure effects in terms of standard deviations of incidence.
We find that years with high levels of antigenic drift coming into an influenza season, e.g.\ drift in A/H3N2 from 2000 to 2001, show high levels of incidence in that season, e.g.\ incidence of A/H3N2 in the 2001/2002 season (Figure~\ref{incidence}A). 
Here, within-clade drift explains 35\% of the variance in standardized incidence and presents a significant correlation (Pearson correlation, $r=0.60$, $p=1.6 \times 10^{-5}$).

We further show that incidence within a clade decreases with antigenic drift in other clades, e.g.\ drift in A/H1N1, B/Vic and B/Yam from 2000 to 2001 decrease incidence of A/H3N2 in the 2001/2002 season (Figure~\ref{incidence}B).
This also represents a significant relationship (Pearson correlation, $r=-0.20$, $p=0.024$).
We fit a linear model that predicts standardized incidence based upon drift within a clade, drift in other clades of same type, and drift in clades of differing type (Table~\ref{interferencetable}).
For example, standardized incidence in A/H3N2 would be explained by drift within A/H3N2, drift in A/H1N1 and by drift across types in B/Vic and B/Yam.
We compared adjusted $R^2$ across models with different regression coefficients included, and arrived at a best-fitting model that explains 46\% of the variance in standardized incidence and includes the terms $\bwithin$, $\bsister$ and $\bother$, but does not include an intercept term.
We find that 1 unit of antigenic drift within a clade increases incidence by 1.26 standard deviations, while 1 unit of drift in a clade of the same type decreases incidence by 0.61 standard deviations and 1 unit of drift in clades of differing type decreases incidence by 0.32 standard deviations.

>>>>

We follow-up this analysis with a more detailed analysis of year-to-year variation in antigenic drift and clade-specific incidence in the USA.
We measure year-to-year antigenic drift within each clade as described in the previous section and shown in Figure \ref{jumps}.
We measure seasonal incidence of each clade by taking average influenza-like illness (ILI) percentage and multiplying this by the proportion of viruses attributable to a clade for each season.
This measure of incidence has previously been shown to have have predictive power in the analysis of seasonal influenza trends \cite{Goldstein11}.
In correlating antigenic drift to seasonal incidence we standardize this measure of incidence across seasons and across clades to measure effects in terms of standard deviations of incidence.
We find that years with high levels of antigenic drift coming into an influenza season, e.g.\ drift in A/H3N2 from 2000 to 2001, show high levels of incidence in that season, e.g.\ incidence of A/H3N2 in the 2001/2002 season (Figure~\ref{incidence}A). 
Here, within-clade drift explains 35\% of the variance in standardized incidence and represents a significant correlation (Pearson correlation, $r=0.60$, $p=1.6 \times 10^{-5}$).

We further show that incidence within a clade decreases with antigenic drift in other clades, e.g.\ drift in A/H1N1, B/Vic and B/Yam from 2000 to 2001 decrease incidence of A/H3N2 in the 2001/2002 season (Figure~\ref{incidence}B).
This also represents a significant relationship (Pearson correlation, $r=-0.20$, $p=0.024$).
We fit a linear model that predicts standardized incidence based upon drift within a clade, drift in other clades of the same virus type, and drift in clades of differing type (Table~\ref{interferencetable}).
For example, standardized incidence in A/H3N2 would be explained by drift within A/H3N2, drift in A/H1N1 and by drift across types in B/Vic and B/Yam.
We compared adjusted $R^2$ across models with different regression coefficients included, and arrived at a best-fitting model that explains 46\% of the variance in standardized incidence and includes the terms $\bwithin$, $\bsister$ and $\bother$, but does not include an intercept term.
We find that 1 unit of antigenic drift within a clade increases incidence by 1.26 standard deviations, while 1 unit of drift in a clade of the same type decreases incidence by 0.61 standard deviations and 1 unit of drift in clades of differing type decreases incidence by 0.32 standard deviations.
